### 2021-01-03 paper/localpsf.tex
moved some h-matrix stuff to the appendix

In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more [CITES]. We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead. The basic ideas behind H-matrices are as follows:
\begin{itemize} 
	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast. Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(n \log^a n)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
\end{itemize}
For more details on H-matrices, we recommend [CITES].

We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE]. Briefly, the H-matrix construction process that we use proceeds as follows. First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}). Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}). Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values. The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries. The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}. In the remainder of this section we describe each of these steps in more detail.

\subsubsection{Cluster trees}
\label{sec:cluster_trees}

\subsubsection{Block cluster tree} 
\label{sec:block_cluster_tree}

\subsubsection{Adaptive cross low rank approximation of blocks}
\label{sec:adaptive_cross}

@@

>>>>

In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more [CITES]. We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead. The basic ideas behind H-matrices are as follows:
\begin{itemize} 
	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast. Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
\end{itemize}
For more details on H-matrices, we recommend [CITES].

We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE]. Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathcal{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small. Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks. The small high rank blocks are formed directly by evaluating their entries. The primary cost of constructing the approximation is the evaluation of $O(N \log N)$ matrix entries of $\widetilde{\mathcal{A}}$, which we perform using Algorithm \ref{alg:eval_y_x}. We describe this H-matrix construction method in detail in Appendix \ref{app:h_matrix}.

@@

\section{Hierarchical matrix construction}
\label{app:h_matrix}

The H-matrix construction process that we use proceeds as follows. First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}). Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}). Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values. The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries. The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.

\subsection{Cluster trees}
\label{sec:cluster_trees}

\subsection{Block cluster tree} 
\label{sec:block_cluster_tree}

\subsection{Adaptive cross low rank approximation of blocks}
\label{sec:adaptive_cross}
